Investigating the alternative prey hypothesis with the POMP framework

### 2024-04-16 workspace/heyuan.Rmd
Add files via upload

```{r}
library(tidyverse)
library(tseries)
```

@@

## POMP

>>>>

```{r}
library(tidyverse)
library(tseries)
#install.packages("ggfortify")
library(ggfortify)
```

@@

ARMA(0,5) Diagnosis
```{r}
fit1 <- arima(ts_cpue, order=c(0,0,5))
residuals <- residuals(fit1)

# ACF check
acf(residuals, main="ACF of Residuals")

# Normality check with Q-Q plot
qqnorm(residuals)
qqline(residuals, col="red")

# Shapiro-Wilk test
shapiro.test(residuals)

# Root Check
ma_coefficients <- coef(fit1)[grepl("ma", names(coef(fit1)))]
MA_roots <- polyroot(c(1, ma_coefficients))
#invertible <- all(Mod(MA_roots) > 1)
#print(paste("Is the model invertible? ", invertible))

autoplot(fit1) +
  ggtitle("ARMA(0, 5)")
```

For the ACF plot of the residuals, all autocorrelations are within the confidence bounds (indicated by blue dashed lines), which suggests that there is no significant autocorrelation present. Additionally, the QQ plot shows that the data points roughly align with the reference line (colored red), supporting the assumption of normality. This is further confirmed by a Shapiro-Wilk test for normality, which returns a p-value of 0.62, indicating that the residuals can be considered normally distributed. Moreover, the roots of the characteristic polynomial are outside the unit circle, confirming that the MA(5) process is invertible—a critical condition for model adequacy. These diagnostics collectively verify that the ARIMA(0,1,5) model is well-specified and adheres to the necessary assumptions for valid time series analysis.

## POMP 1

```{r}
install.packages("pomp")
library(pomp)
```

To create a POMP for CPUE, I need to define the state process that reflects the underlying dynamics of the species population and the observation process that relates the state to the CPUE measurements.
